Switch equivalence and weak prime domains for fusions - quotients

### 2024-08-15 quozienti.tex
inserite unioni e parte sui sistemi di fattorizzazioni

@@

%identità
\newcommand{\id}[1]{\mathsf{id}_{#1}}
%codominio
\newcommand{\cod}[1]{\mathsf{cod}({#1})}
	%Variabili categorie
\def\A{\textbf {\textup{A}}}
%mono
\newcommand{\mto}[0]{\scalebox{1}{$\rightarrowtail$}}

@@

\def\R{\mathsf{R}}
\def\B{\textbf {\textup{B}}}
\def\C{\textbf {\textup{C}}}
\def\D{\textbf {\textup{D}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\def\E{\textbf {\textup{E}}}

@@

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{notation}{Notation}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\title{Switch equivalence and weak prime domains for fusions}
\author{Paolo Baldan\and Davide Castelnovo\and  Andrea Corradini\and Fabio Gadducci}

@@

\subsection{$\mathcal{M}$-subobjects and factorization systems}\label{app:sub}

@@

Now, let us recall the following lemma, which is the application of the Adjoint Functor Theorem \cite{freyd1964abelian,freyd1990categories} to the partial ordered case.

@@

The previous Lemma, together with \Cref{cor:fact} and \Cref{prop:finlim}, allows us to deduce the following result.

@@

\section{$\mathcal{M}$-adhesivity is not enough}\label{app:fill}
 In \Cref{prop:equi} we proved that, in the linear case, the existence of an indpendence pair between two derivation is equivalent to that of a filler between them. This result can be further refined: in a\cite{baldan2011adhesivity} a class  $\mathbb{B}$ of (quasi)adhesive category is defined for which the local Church-Rosser Theorem holds even for left-linear DPO-rewriting system. In our language, and given \Cref{prop:fil} and \Cref{rem:locCR}, this amount to prove that, for elements of $\mathbb{B}$, every independence pair induces a filler.

\begin{definition}Let $\X$ be a category, we say that $\X$ satisfies
	\begin{itemize}
		\item the \emph{mixed decomposition} property if for every diagram
		\[\xymatrix{X \ar[d]_{a} \ar[r]^{f}& \ar[r]^{g} Y \ar[d]^{b}& Z \ar[d]^{c}\\ A \ar[r]_{h}& B \ar[r]_{k}& C}\]
		whose outer boundary is a pushout and in which $k$ is a monomorphisms, 
		\item the \emph{pushout decomposition} property
	\end{itemize}
\end{definition}

\begin{lemma}
	contenuto...
\end{lemma}
\begin{proof}
	contenuto...\qedhere 
\end{proof}

\begin{corollary}
	contenuto...
\end{corollary}

The following result shows that the mixed and pushout decomposition properties guarantee that every independence pair gives rise to a filler.

\begin{theorem}
	\todo{filler e classe B+}
\end{theorem}
\begin{proof}\qedhere 
\end{proof}

Our next step is to identify sufficient conditions for a category $\X$ to satisfy the mixed and pushout decomposition properties.

\begin{definition}
	\todo{classe B e class B+}
\end{definition}

\begin{example}
\todo{esempi}
\end{example}

\begin{example}
	\todo{esempi}
\end{example}

\begin{proposition}
	\todo{da B a B+}
\end{proposition}
\begin{proof}
	contenuto... \qedhere 
\end{proof}
\begin{lemma}\todo{due proprietà classe B}
\end{lemma}
\begin{proof}
	\qedhere 
\end{proof}

\begin{corollary}
	\todo{due proprietà classe B+}
\end{corollary}

\begin{corollary}
	\todo{filler e classe B+}
\end{corollary}

\section{Rewriting systems on $\Set$.}\label{app:set}

\section{On permutations}\label{app:perm}
\todo{inserire risutati sulle permutazione che abbiamo usato}

\begin{definition}
	\todo{inversioni}
\end{definition}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

\newcommand{\cmd}{\mathbf{IC}}

\newcommand{\cori}{\mathbf{CR}}

@@

%identità
\newcommand{\id}[1]{\mathsf{id}_{#1}}
%codominio
\newcommand{\cod}[1]{\mathsf{cod}({#1})}
	%Variabili categorie
\def\A{\textbf {\textup{A}}}
%arrows
\newcommand{\rightarrowdbl}{\rightarrow\mathrel{\mkern-14mu}\rightarrow}

\newcommand{\xrightarrowdbl}[2][]{%
	\xrightarrow[#1]{#2}\mathrel{\mkern-14mu}\rightarrow
}

\newcommand{\mini}[0]{\scalebox{.4}{$\bullet$}}
\newcommand{\mto}[0]{\rightarrowtail}
\newcommand{\rto}[0]{\xrightarrow{\mini}}
\newcommand{\fto}[0]{\xrightarrowdbl{\mini}}
\newcommand{\eto}[0]{\twoheadrightarrow}

@@

\def\R{\mathsf{R}}
\def\B{\textbf {\textup{B}}}
\def\C{\textbf {\textup{C}}}
\def\D{\textbf {\textup{D}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\def\Z{\textbf {\textup{Z}}}
\def\E{\textbf {\textup{E}}}
\newcommand{\m}[1]{\mathbb{#1}}

@@

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{notation}{Notation}
\newtheorem*{setting}{Setting}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\title{Switch equivalence and weak prime domains for fusions - quotients}
\author{Paolo Baldan\and Davide Castelnovo\and  Andrea Corradini\and Fabio Gadducci}

@@

\subsection{$\mathcal{M}$-subobjects and factorization systems}

@@

Now, let us recall the following lemma, which is the application of the Adjoint Functor Theorem to the partial ordered case.

@@

The previous result, together with \Cref{cor:fact} and \Cref{prop:finlim}, allows us to deduce the following result, which has been used in \cite{baldan2011lattice,braatz2010finitary} to show that certain $\mathcal{M}$-adhesive categories have a factorization system.

@@

\section{Idempotent comonads}

\begin{definition}
	A \emph{comonad} $\m{C}$ on a category $\X$ is a triple $(C, \delta, \epsilon)$ given by a functor $C\colon \X\to \X$ and two natural transformations $\delta\colon C\to C\circ C $, $\epsilon\colon C\to \id{\X}$, called, respectively \emph{comultiplication} and \emph{counit} such that the following diagram commute
	\[\xymatrix@C=30pt{C \ar[d]_{\delta}\ar[r]^-{\delta}& C\circ C  \ar[d]^{\delta*C}&  C \ar[d]_{\delta}\ar[r]^{\delta} \ar[dr]^-{\id{C}} & C\circ C \ar[d]^{\epsilon *C}\\ C\circ C \ar[r]_-{C*\delta}& C\circ C\circ C & C\circ C \ar[r]_-{C*\epsilon} & C}\]

	A comonad $\m{C}$ is \emph{idempotent} if $\delta$ is an isomorphism.  A \emph{morphism} between two comonad $\m{C}=(C, \delta, \epsilon)$ and $\m{D}=(D, \mu, \eta)$ is a natural transformation $\chi\colon C\to D$ such that the following diagrams commute
	\[\xymatrix@C=30pt{ C \ar[r]^-{\delta} \ar[d]_{\chi} & C\circ C \ar[d]^{\chi*\chi}& C \ar[r]^{\epsilon} \ar[d]_{\chi}& 
\id{\X}  \\ D \ar[r]_-{\mu} & D\circ D
	& D \ar[ur]_{\eta} }\]

	We will denote by $\cmd(\X)$ the (possibly not locally small) category whose objects are idempotent comonads and whose arrows are morphisms of monads.
\end{definition}

\begin{remark} Let us recall that, given two functors $F, G\colon \X\rightrightarrows \Y$ and a natural transformation $\chi$ between them, then for every $X\in \X$ we have a naturality square
		\[\xymatrix@C=40pt{F(F(X)) \ar[r]^{F(\chi_X)} \ar[d]_{\chi_{F(X)}}& F(G(X)) \ar[d]^{\chi_{G(X)}}\\G(F(X)) \ar[r]_{G(\chi_X)} & G(G(X))}\]

	This entails that $(\chi * G)\circ (F*\chi)$ and $(G*\chi) \circ (\chi* F)$ are the same natural transformation, which we denote by $\chi*\chi$.
\end{remark}

\begin{example}\label{ex:1}
	\todo{connessi}
\end{example}

\begin{example}\label{ex:2}
	\todo{grafi con equivalenza}
\end{example}
\begin{example}
	\todo{Ce ne sono altri? sarebbe bello usare i calcoli ma credo rientrino nel primo}
\end{example}

We recall the following lemma relating idempotent comonads and coreflective subcategories.

\begin{lemma}\label{lem:comsub}
	Let $L\colon \Y\to \X$ and $R\colon \X \to \Y$ be two functors such that $L\dashv R$
\end{lemma}

Associated to any idempotent comonad we have a class of arrows.

\begin{definition}
	Let $\m{T}$ be an idempotent comonad on a category $\X$. We say that an arrow $f:X\to Y$ i
\end{definition}
\begin{remark}
	contenuto...
\end{remark}

\subsection{A left adjoint to $\mathcal{R}_{(-)}$}

\begin{setting}Let $\X$ be a category with a proper factorization system $(\mathcal{E}, \mathcal{M})$. Our abstract setting to deal with fusions begins with a coreflective subcategory $I\colon \Y\to \X$ with coreflector $F\colon \X\to \Y$ with the following properties:
	\begin{enumerate}
		\item the counit $\epsilon \colon I\circ F\to \id{\X}$ has all components in $\mathcal{M}$; 
		\item given a pushout as the one below, if $F(f)$ is mono then $F(h)$ is mono too.
		\[\xymatrix{X \ar[r]^{g} \ar[d]_{f}& Z \ar[d]^{h}\\ Y\ar[r]_{k} & W}\]
	\end{enumerate}

We denote by $\mathcal{R}$ the class of all arrows of $\X$ whose image in $\Y$ is mono.  A \emph{fusion} is an arrow belonging to both $\mathcal{R}$ and $\mathcal{E}$.
 \end{setting}
 \begin{notation}
 	Given an arrow $f\colon X\to Y$ of $\X$, we will use the following notational conventions:
 	\begin{itemize}
 		\item $f\colon X\mto Y$ if $f$ belongs to $\mathcal{M}$;
 		\item $f\colon X\eto Y$ if $f$ belongs to $\mathcal{E}$;
 		\item $f\colon X \rto Y$ if $f$ belongs to $\mathcal{R}$;
 		\item $f\colon X \fto Y$ if $f$ is a fusion.
 	\end{itemize}
 \end{notation}

 \begin{remark}The following facts hold true:
 	\begin{itemize}
 		\item since $F$ us a right adjoint, it sends monos to monos, thus all monos, in particular those in $\mathcal{M}$, are in $\mathcal{R}$;
 		\item $\mathcal{R}$ is closed under composition and pullbacks;
 		\item if $g\circ f$ is in $\mathcal{R}$, then $f$ is in $\mathcal{R}$ too.
 	\end{itemize}
 \end{remark}

The main property in which we are interested in is given by the following proposition.

\begin{proposition}\label{prop:main}
	Let $h\colon X \rto Y$ be an arrow in $\mathcal{R}$. Given an object $L$ of $\Y$ and $f,g\colon L\rightrightarrows X$, if $h\circ g=h\circ f$ then $g=f$.
\end{proposition}
\begin{proof}
	Since $h\circ g=h\circ f$ and $h$ is in $\mathcal{R}$ we have that $F(g)=F(h)$. Thus we have:
	\begin{align*}
		f\circ \epsilon_{L}&=\epsilon_X\circ I(F(f))\\&=\epsilon_X\circ I(F(g))\\&=g\circ \epsilon_L
	\end{align*}
	By hypothesis $L$ is in $\Y$, thus $\epsilon_L$ is an isomorphism and we can conclude.
\end{proof}

We can wonder if, given a class of arrows in a sufficiently nice category $\X$, we can recover all the properties of our setting. This is the content of the lemma below. We need a technical definition first.

\begin{definition}
	Let $\mathcal{M}$ be a class of monos in $\X$. We say that $\mathcal{M}$ has \emph{$\mathcal{M}$-unions} if, for every object $X$,  $\sub{M}{\X}{X}$ has all suprema and these are computed by multiple pushouts. 
\end{definition}

\begin{remark}
	Being computed by multiple pushout means that the supremum of a family $\{[m_j]\}_{j\in J}$ in $\sub{M}{\X}{X}$ is represented by an arrow $m\colon M\to X$ with the following universal property: given the solid part of the diagram below, there exists a unique dotted arrow $t\colon M\to Z$ making it commutative.
	\[\xymatrix{& M_j \ar[d]_{k_j} \ar@{>->}[dr]^{m_j} \\Z  \ar@{<-}[ur]^{h_j} \ar@{>->}@/_.4cm/[rr]_{z} \ar@{<.}[r]^t & M \ar@{>->}[r]^m & X }\]

	In particular, if $J$ is empty, we get an arrow $o_X\colon O_X\to X$ which is initial in $\mathcal{M}/X$. Notice that, in particular, $o_X$ is the unique element of $\X(O_X, X)$. Indeed, if $f$ is another arrow $O_X\to X$ then we can factor it as $m\circ e$ with $m$ in $\mathcal{M}$. Thus there 
\end{remark}

\begin{lemma}	
Let $\X$ be an $\mathcal{M}$-adhesive category $\X$  with a proper and stable factorization system $(\mathcal{E}, \mathcal{M})$. Suppose that $\mathcal{M}$ has $\mathcal{M}$-unions. Then there exists a corefelctive subcategory $I\colon \Y\to \X$ satisfying the assumptions of our setting.
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}

\section{Rewriting with fusions}

 \begin{setting}
 	contenuto...
 \end{setting}

\begin{definition}

\end{definition}

\bibliographystyle{plain}
\bibliography{bibliog.bib}
\end{document}
